### 2017-10-15 mcmc.tex
tweaks

One thing we can keep in mind to guide us in this is what we said above (\sectionname~\ref{sec:sampling}),
  which is that samplings are good for doing \emph{integrals}.
We should endeavor to use as our ``results'' outputs from the \MCMC\
  that are based on integrals computed with the sampling.
This includes expectations, medians, quantiles, one-dimensional histograms,
  and multi-dimensional histograms.
This does \emph{not} include the ``best'' sample or a mode or optimum.
The latter things are not necessarily illegitimate outputs of the \MCMC,
  but they do not make best use of the fact that the sampling is a tool for integration,
  and we do not recommend them.

@@

For example,
  imagine that by \MCMC\ you have generated
  $K$ samples $\pars_k$ of a parameter vector $\pars$.
Now you have a scalar function $g(\pars)$ which takes the parameter vector
  and returns a scalar value%
  \note{For the median-of-sampling value, $g(\pars)$ needs to return a scalar,
  but technically, for the mean-of-sampling, $g(\pars)$ can be a vector or something high-dimensional.}
  which could be as simple as a single component of the parameter vector
  (one parameter from the list),
  or something more complex.
The mean-of-sampling value $\mean{q}$ is just
\begin{eqnarray}
\mean{q} &\leftarrow& \frac{1}{K}\,\sum_{k=1}^K g(\pars_k)
\quad ,
\end{eqnarray}
  and the median-of-sampling value is just the $[K/2]$th value of $g(\pars_k)$
  when the $g(\pars_k)$ have been ordered (sorted).
These are both produced by integrals;
  the first is the value returned by the expectation integral estimate,
  the second is the value past which half of the integrated pdf lies.

@@

As we mentioned above, scientists often like to report the ``best sample'';
  that is, the sample with the highest posterior pdf value.
This is not usually a good idea.
For one, \MCMC\ is a sampling algorithm, not an optimization algorithm;
  there are no guarantees that it will find the optimum of the posterior pdf in reasonable time.%
  \note{The sampler samples the pdf fairly.
    The tallest peak in the posterior is not guaranteed to be---%
    and in general is not---%
    also the peak with the greatest posterior ``mass''.
    That is, the samples in a fair sampling will not necessarily come
    predominantly from the \emph{tallest} peak in the posterior pdf.
    They will come from the peak with the greatest total integral
    of the likelihood over the prior.
    These issues might sound ``academic'',
    but when the number of parameters gets large (greater than, say, 10),
    many normal intuitions one might have (if any) about what is reasonably likely in a pdf
    are regularly violated.}
For two, the closeness of the best sample to the posterior mode is a strong function of sample size,
  and with very bad scaling.%
  \note{If you have a $K$-point sampling in a $D$-dimensional parameter space,
    there is some best sample $\pars_k^{\best}$.  Now imagine that you want to make
    that best sample 10 times better---that is, you want to find a point $\pars^{\better}$
    that is ten times closer to the true optimum.  How many more samples do you need to take?
    In the limit that $K$ is large, on average, you have to take a factor of
    $10^D$ more samples!
    Any reasonable optimizer is far, far better than this (in terms of compute time).}
For three, if you just want an optimum, use an optimizer!
That is, we \emph{strongly advise against} reporting,
  as ``the'' measurement or ``the result'',
  the parameter value or values for the best sample.
Giving some posterior samples is a good idea (we return to this below).

@@

If you \emph{do} return posterior samples to your audience, how many
should you return?
There are only heuristic answers to this, but a guideline is that, if
you only care about a single (one-dimensional) parameter, you don't
need many samples; a dozen suffices to give you a reasonable posterior
pdf mean and variance.\note{This point is made well by \citet{mackay}.}
If you care about $K$ parameters or dimensions of the parameter space,
in general the need for samples grows exponentially (or perhaps even
factorially!)\ with the number of dimensions.
So if you have a ten-dimensional space, expect to be publishing your
samples in an electronic table on the web!

@@

\begin{itemize}

\item \emph{Trace plots}~---~The first set of plots are the ``trace
plots''~--~the parameter values as a function of step number.
These plots can be used to select burn-in lengths, indicate problems with the
model or sampler, and qualitatively judge convergence.
This being said, it is important to remember that \emph{good looking trace
plots do not guarantee converged sampling} and, even if the traces look okay,
the convergence diagnostics discussed in \sectionname~\ref{sec:convergence}
might identify problems.

>>>>

One thing we can keep in mind to guide us in this is what we said above
  (in \sectionname~\ref{sec:sampling}),
  which is that samplings are good for doing \emph{integrals}.
We should endeavor to use as our ``results'' outputs from the \MCMC\
  that are based on integrals computed with the sampling.
This includes expectations, medians, quantiles, one-dimensional histograms,
  and multi-dimensional histograms.
This does \emph{not} include the ``best'' sample or a mode or optimum.
The latter things are not necessarily illegitimate outputs of the \MCMC,
  but they do not make best use of the fact that the sampling is a tool for integration,
  and we do not recommend them.

@@

For example,
  imagine that by \MCMC\ you have generated
  $K$ samples $\pars_k$ of a parameter vector $\pars$.
Now you have a scalar function $g(\pars)$ which takes the parameter vector
  and returns a scalar value%
  \note{For the median-of-sampling value, $g(\pars)$ needs to return a scalar,
  but technically, for the mean-of-sampling, $g(\pars)$ can be a vector or something high-dimensional.}
  which could be as simple as a single component of the parameter vector
  (one parameter from the list),
  or something more complex.
The mean-of-sampling value $\mean{q}$ is just
\begin{eqnarray}
\mean{q} &\leftarrow& \frac{1}{K}\,\sum_{k=1}^K g(\pars_k)
\quad ,
\end{eqnarray}
  and the median-of-sampling value is just the $[K/2]$th value of $g(\pars_k)$
  when the $g(\pars_k)$ have been ordered (sorted) from lowest to highest.
These are both produced by integrals;
  the first is the value returned by the expectation integral estimate,
  the second is the value past which half of the integrated pdf lies.

@@

As we mentioned above, scientists often like to report the ``best sample'';
  that is, the sample with the highest posterior pdf or $f(\pars)$ value.
This is not usually a good idea.
For one, \MCMC\ is a sampling algorithm, not an optimization algorithm;
  there are no guarantees that it will find the optimum of the posterior pdf in reasonable time.%
  \note{The sampler samples the pdf fairly.
    The tallest peak in the posterior is not guaranteed to be---%
    and in general is not---%
    also the peak with the greatest posterior ``mass''.
    That is, the samples in a fair sampling will not necessarily come
    predominantly from the \emph{tallest} peak in the posterior pdf.
    They will come from the peak with the greatest total integral
    of the likelihood over the prior.
    These issues might sound ``academic'',
    but when the number of parameters gets large (greater than, say, 10),
    many normal intuitions one might have (if any) about what is reasonably likely in a pdf
    are regularly violated.}
For two, the closeness of the best sample to the posterior mode is a strong function of sample size,
  and with very bad scaling.%
  \note{If you have a $K$-point sampling in a $D$-dimensional parameter space,
    there is some best sample $\pars_k^{\best}$.  Now imagine that you want to make
    that best sample 10 times better---that is, you want to find a point $\pars^{\better}$
    that is ten times closer to the true optimum.  How many more samples do you need to take?
    In the limit that $K$ is large, on average, you have to take a factor of
    $10^D$ more samples!
    Any reasonable optimizer is far, far better than this (in terms of compute time).}
For three, if you just want an optimum, use an optimizer!
That is, we \emph{strongly advise against} reporting,
  as ``the'' measurement or ``the result'',
  the parameter value or values for the best sample.
Giving some posterior samples is a good idea (we return to this below).

@@

If you \emph{do} return posterior samples to your audience, how many
should you return?
There are only heuristic answers to this, but a guideline is that, if
you only care about a single (one-dimensional) parameter, you don't
need many samples; a dozen suffices to give you a reasonable posterior
pdf mean and variance.\note{This point is made well by \citet{mackay}.}
If you care about $K$ parameters or dimensions of the parameter space,
in general the need for samples grows exponentially (or perhaps even
factorially!) with the number of dimensions.
So if you have a ten-dimensional space, expect to be publishing your
samples in an electronic table on the web!

@@

\begin{itemize}
\item \emph{Trace plots}~---~The first set of plots are the ``trace
plots''~--~the parameter values as a function of step number.
These plots can be used to select burn-in lengths, indicate problems with the
model or sampler, and qualitatively judge convergence.
This being said, it is important to remember that \emph{good looking trace
plots do not guarantee converged sampling} and, even if the traces look okay,
the convergence diagnostics discussed in \sectionname~\ref{sec:convergence}
might identify problems.
